---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
assignurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(assignurl, destfile = "activity.zip", method = "curl")
unzip("activity.zip")
assigndata <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
```

## What is mean total number of steps taken per day?
Calculating the total number of steps taken per day
```{r}
data_per_date <- assigndata %>% select(date, steps) %>% group_by(date) %>% summarize(totalsteps = sum(steps)) %>% na.omit()
head(data_per_date)
```

Making a histogram of the total number of steps taken per day
```{r}
hist(data_per_date$totalsteps, xlab = "Total Daily Steps", ylab = "Frequency", main = "Histogram of Total Steps Taken per Day", breaks = 20)
```

Calculating the mean and median of the total number of steps taken per day
```{r}
mean(data_per_date$totalsteps)
median(data_per_date$totalsteps)
```


## What is the average daily activity pattern?
Making a time series plot of interval and average steps taken across all days
```{r}
data_per_interval <- assigndata %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(meansteps = mean(steps))

ggplot(data_per_interval, aes(x = interval, y = meansteps)) + geom_line() + labs ( x = "Interval of Steps Taken") + labs (y = "Average Steps Taken") + labs (title = "Time-series plot of Interval and Average Steps Taken across all days")
```

Identifying the interval with maximum number of steps
```{r}
data_per_interval[which(data_per_interval$meansteps == max(data_per_interval$meansteps)),]
```


## Imputing missing values
Calculating the total number of missing values
```{r}
sapply(X = assigndata, FUN = function(x) sum(is.na(x)))
```

Replacing all missing values with the mean of that interval
```{r}
replace_with_mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
```

Creating new dataset with replaced missing values
```{r}
replaced_data <- assigndata %>% group_by(interval) %>% mutate(steps = replace_with_mean(steps))
```

Making a histogram of total number of steps taken each day with new replaced data
```{r}
replaced_data_per_date <- replaced_data %>% select(date, steps) %>% group_by(date) %>% summarize(totalsteps = sum(steps))
hist(replaced_data_per_date$totalsteps, xlab = "Total Daily Steps", ylab = "Frequency", main = "Histogram of Total Steps Taken per Day", breaks = 20)
```

Calculating mean and median of total steps taken each day of replaced data
```{r}
mean(replaced_data_per_date$totalsteps)
median(replaced_data_per_date$totalsteps)
```

Finding the diffeence in mean and median between original and replaced data
```{r}
mean(replaced_data_per_date$totalsteps) - mean(data_per_date$totalsteps)
median(replaced_data_per_date$totalsteps) - median(data_per_date$totalsteps)
```


## Are there differences in activity patterns between weekdays and weekends?
Creating a factor variable to categorize date into weekday and weekend
```{r}
replaced_data$weekday_weekend <- ifelse(weekdays(as.Date(replaced_data$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "Weekday", "Weekend")
```

Making a panel plot to compare interval and average steps taken by weekdays
```{r}
replaced_data_per_interval <- replaced_data %>% group_by(interval, weekday_weekend) %>% summarize(meansteps = mean(steps))
ggplot(replaced_data_per_interval, aes(x = interval, y = meansteps, color = weekday_weekend)) + geom_line() + facet_grid(weekday_weekend~.) + labs ( x = "Interval of Steps Taken") + labs (y = "Average Steps Taken") + labs (title = "Comparison of Intervals and Average Steps Taken by weekdays")
```
